---
title: Louisville's Heroin Epidemic
author: Adam
date: '2017-08-18'
slug: louisville-s-heroin-epidemic
categories:
  - City Development
  - Crime
  - Drugs
tags:
  - Mapping
  - Visualization
  - R
  - Drugs
---


If you have been paying any attention to the news recently, you will have heard
that the United States is experiencing a resurgance of heroin use.  Unfortunately, 
Louisville is smack in the middle of this growing epidemic. Locally, I think the perception
is that one, this problem sprung up overnight with no warning and two, even if heroin use *is* a problem, it isn't
a problem near *me*.  But is there truth to either of these beliefs?  

Being the data geek that I am, these questions piqued my curiosity. So instead of just accepting these reports as fact,
I decided to see just how much of a problem heroin actually is in our region.  Using heroin crime data 
(namely possession and trafficking) from Louisville's open crime dataset as a proxy for use, I decided to 
visualize Louisville's heroin problem to see if I could gain any insight.  Is there truly a heroin epidemic? 
How fast did this become a problem? And probably most important to the average person, is this actually happening near me? 
Sadly, the answers appear to be 'Yes', 'Quickly', and 'Absolutely'. 

```{r package_load, cache = TRUE, echo = FALSE, message=FALSE, warning = FALSE}
#Loading all packages necessary for report
library(stringi)
library(dplyr)
library(ggplot2)
library(readr)
library(viridis)
library(gridExtra)
library(tidyr)
library(rgeos)
library(rgdal)
library(raster)
library(broom)
library(gtable)
library(grid)
```

### Overview of Various Drug Offenses


```{r, load_data_and_base_filter, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE}
# Original dataset is found here (http://portal.louisvilleky.gov/dataset/crimedataall-data)
# Geocoding and intial cleaning/formatting process can be found at
# https://github.com/Grollus/LouisvilleCrime/blob/master/louisville_raw_data_and_geocoding.R


# Since most data cleaning was done outside this report, all I have to do here is
# create any useful additional variables and load the file
raw_data <- read_rds("lou_crime_geocoded_df.rds.gzip")

create_date_variables <- function(df){
  require(lubridate)
  # Uses the POSIXct date_occured variable to create useful date related
  # subvariables
  df$year <- year(df$date_occured)
  df$month <- month(df$date_occured)
  df$day <- day(df$date_occured)
  df$hour <- hour(df$date_occured)
  df$year_month <- paste(df$year, df$month, sep = '-')
  df$day_of_week <- wday(df$date_occured, label = TRUE, abbr = FALSE)
  df$weekday <- ifelse(df$day_of_week == "Saturday" | df$day_of_week == "Sunday", 
                              "Weekend", "Weekday")
  df$yday <- yday(df$date_occured)
  
  return(df)
}

crime_lou <- create_date_variables(raw_data)
# Filter out records with no lat/lng coords and those from desired time period
crime_lou <- crime_lou%>%
  filter(year <=2016 & year >=2005 & !is.na(lat)  & !is.na(lng))
```


```{r, theme_and_drug_filter, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE}
# Base theme for all bar charts
bar.theme <- 
  theme(legend.position = "none")+
  theme(plot.title = element_text(size = 18, family = "serif", face = 'bold', hjust = 0, color = "#666666"))+
  theme(axis.title = element_text(size = 14, family = "serif", face = 'bold', hjust = 0, color = "#666666"))+
  theme(axis.title.y = element_text(angle = 90))+
  theme(axis.title.x = element_text(hjust = 0, vjust = 0.5))+
  theme_bw()


########################################################################################################
# Going to regex out all of the drug offenses I'm interested in. Then I can use this for easy faceting
# and just analysis in general. I create a new variable named 'drug' that just identifies the 
# associated drug with that crime entry.

drug_offenses <- crime_lou%>%
  filter(stri_detect_regex(uor_desc, "heroin|opiate|cocaine|marijuana|meth|codeine|lsd|barbituate|pcp|
                           hallucinogen"))%>%
  mutate(drug = ifelse(stri_detect_regex(uor_desc, "heroin") == TRUE, "Heroin", 
                       ifelse(stri_detect_regex(uor_desc, "opiate|codeine") == TRUE, 'Opiates', 
                              ifelse(stri_detect_regex(uor_desc, "meth") == TRUE, "Methamphetamine",
                                     ifelse(stri_detect_regex(uor_desc, "cocaine") == TRUE, "Cocaine",
                                            ifelse(stri_detect_regex(uor_desc, "marijuana") == TRUE, "Marijuana", "Other"))))))


# take input drug and two years; returns a df with the count of offenses
# for input drug for those years
delta_between_two_years <- function(which_drug, y1, y2){
  drug_offenses%>%
    filter(drug == which_drug, year == y1|year == y2)%>%
    group_by(year)%>%
    summarise(count = n())
}


per_change_heroin <- delta_between_two_years("Heroin", 2007, 2016)
per_change_meth <- delta_between_two_years("Methamphetamine", 2007, 2016)
per_change_opiates <- delta_between_two_years("Opiates", 2007, 2016)
per_change_cocaine <- delta_between_two_years("Cocaine", 2007, 2016)
per_change_marijuana <- delta_between_two_years("Marijuana", 2007, 2016)

```

To get a sense of the general drug 'scene' in Louisville I want to look at the number
of offenses of the major drugs over the past decade.  I can't really tell if heroin is a problem if
I have no idea of the general level of drug use in our city.  

Using some simple bar plots, the huge increase in heroin offenses is immediately obvious. We have seen 
a staggering *`r round(per_change_heroin$count[2]/per_change_heroin$count[1]* 100, 0)`%* increase in heroin 
offenses from 2007 to 2016.  While it is true we have seen an increase in methamphetamine and opiate offenses--by 
*`r round(per_change_meth$count[2]/per_change_meth$count[1]* 100, 0)`%* and 
*`r round(per_change_opiates$count[2]/per_change_opiates$count[1]* 100, 0)`%* respectively--
it is not to nearly the same degree. Heroin's offenses have increased over 3 times as fast as
those of meth and other opiates, even though we are seeing large increases in crime related
to those drugs too!


```{r, increasing_drugs_bar_plot, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE, out.width = "90%"}
drug_offenses %>%
  filter(year >= 2007, drug != "Marijuana" & drug != "Cocaine" & drug != "Other")%>%
  group_by(year, drug)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  geom_bar(stat = 'identity', fill = "#0072B2", alpha = .5, color = 'black')+
  facet_wrap(~drug)+
  stat_smooth(method = 'auto', size = .45, alpha = .2, fill = 'grey', colour = 'red2', level = .95)+
  ggtitle("Heroin, Meth, and Opiate Offenses, 2007-2016")+
  labs(x = "Year", y = "Offenses")+
  scale_x_discrete(limits = seq(2007, 2016, by = 1))+
  bar.theme+
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.background = element_rect(fill = 'white'),
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 0, vjust = .5, size = 5)
)

```


When I first saw this--after the initial shock--I figured drug use in general over the past
decade must just be up.  A rising tide lifts all boats so to speak.  However, when you plot
out the other major recreational drugs, you don't see this.  We have seen fairly sizeable decreases in criminal activity relating
to cocaine and marijuana(*-`r round(per_change_cocaine$count[2]/per_change_cocaine$count[1]* 100, 0)`%* and 
*-`r round(per_change_marijuana$count[2]/per_change_marijuana$count[1]* 100, 0)`%* respectively)
over the same period of time.  

```{r, decreasing_drugs_bar_plot, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE, out.width = "90%"}
drug_offenses %>%
  filter(year >= 2007, drug == "Marijuana"| drug == "Cocaine")%>%
  group_by(year, drug)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  geom_bar(stat = 'identity', fill = "#0072B2", alpha = .5, color = 'black')+
  facet_wrap(~drug)+
  stat_smooth(method = 'auto', size = .45, alpha = .2, fill = 'grey', colour = 'red2', level = .95)+
  ggtitle("Cocaine and Marijuana Offenses, 2007-2016")+
  labs(x = "Year", y = "Offenses")+
  scale_x_discrete(limits = seq(2007, 2016, by = 1))+
  bar.theme+
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.background = element_rect(fill = 'white'),
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 0, vjust = .5, size = 7)
)

```


In fact, if we look at a plot of all recreational drug offenses, we see a slight trend 
downward over the past decade.

```{r,all_rec_drug_offenses, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE, out.width = "90%"}
drug_offenses %>%
  filter(year >= 2007)%>%
  group_by(year)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  geom_bar(stat = 'identity', fill = "#0072B2", alpha = .5, color = 'black')+
  stat_smooth(method = 'lm', size = .45, alpha = .2, fill = 'grey', colour = 'red2', level = .95)+
  ggtitle("All Recreational Drug Offenses, 2007-2016")+
  labs(x = "Year", y = "Offenses")+
  scale_x_discrete(limits = seq(2007, 2016, by = 1))+
  bar.theme+
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.background = element_rect(fill = 'white'),
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 0, vjust = .5, size = 7)
)

```

```{r diff_mari_heroin, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE}
times_more_mary_v_heroin <- drug_offenses%>%
  filter(drug == 'Heroin'|drug == 'Marijuana', year == 2016)%>%
  group_by(drug)%>%
  summarise(count = n())
```

A couple important points probably should be noted here. First, if you look at the scale of 
the y-axis on all the previous plots, you will notice that marijuana offenses dominate the overall
count. If you plot heroin and marijuana offenses on the same scale, heroin use still appears miniscule.
As of 2016 we still had `r round(times_more_mary_v_heroin$count[2]/times_more_mary_v_heroin$count[1], 1)`
times more marijuana offenses than heroin offenses.


```{r, marijuana_v_heroin_bar, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE, out.width="90%"}
drug_offenses %>%
  filter(year >= 2007, drug == "Marijuana"|drug == "Heroin")%>%
  group_by(year, drug)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  facet_wrap(~drug)+
  geom_bar(stat = 'identity', fill = "#0072B2", alpha = .5, color = 'black')+
  # stat_smooth(method = 'auto', size = .45, alpha = .2, fill = 'grey', colour = 'red2', level = .95)+
  ggtitle("Heroin v Marijuana Offenses, 2007-2016")+
  labs(x = "Year", y = "Offenses")+
  scale_x_discrete(limits = seq(2007, 2016, by = 1))+
  bar.theme+
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.background = element_rect(fill = 'white'),
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 0, vjust = .5, size = 7)
)

```

This is not to say that heroin use is small and inconsequential.  What this means is that marijuana 
offenses have the power to move(and very possibly distort) the entire drug picture by itself.  As can often be the case, taking 
a very aggregated view of the data would lead you to a completely incorrect conclusion.

Secondly, we need to keep in mind that none of this is happening in a vacuum. There are very real social, 
political, and economic changes that happen which *do* end up altering the landscape.  Unfortunately,
these changes can make it very difficult to pin down exactly what is happening. For instance, when Obama
first took office in 2008, he voiced a softer federal stance on prosecuting marijuana crimes.  So
while marijuana was--and still *is*--illegal in all forms in Kentucky, did this federal statement
affect the enforcement and prosecution of marijuana offenses in Louisville? It is way too complicated of an issue 
to answer simply yes or no, but I think it would be naive to think it had *zero* effect.  But we can't be sure
(at least with this simplistic analysis) that there wasn't simply a coincidental decline in marijuana use, which also resulted in fewer
marijuana crimes.  This same sort of logic applies to offenses of all types.

In that vein, we have to at least mention the changes in the opiate drug landscape that could explain 
some of this heroin epidemic.  Back in 2010 and 2011, two of the major opiates abused recreationally
were [reformulated](https://odcp.ky.gov/Pages/The-Heroin-Epidemic.aspx) to make recreational use
more difficult. Around the same time, various systems(for instance KASPER) were put in place to make abusing
prescription opiates more difficult. Though it is always difficult to establish causation, it seems
very likely that effectively restricting the supply of prescription opiates around 2011 triggered
a surge in heroin demand as a replacement.  




```{r, overdose_heroin_correlation, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE}
overdose <- crime_lou%>%
  filter(uor_desc == 'overdose'|stri_detect_regex(uor_desc, "heroin"))%>%
  mutate(Category = ifelse(stri_detect_regex(uor_desc, 'heroin') == TRUE, "Offenses", "Overdoses"))%>%
  group_by(year, Category)%>%
  summarise(count = n())

methoverdose <- crime_lou%>%
  filter(uor_desc == 'overdose'|stri_detect_regex(uor_desc, "meth"))%>%
  mutate(Category = ifelse(stri_detect_regex(uor_desc, 'meth') == TRUE, "Offenses", "Overdoses"))%>%
  group_by(year, Category)%>%
  summarise(count = n())

methcor_df <- methoverdose %>%
  spread(key = year, value = count)%>%
  gather(year, count, -Category)
methcor_df$count <- ifelse(is.na(methcor_df$count) == TRUE, 0, methcor_df$count)

methover <- filter(methcor_df, Category == "Overdoses")
methoff <- filter(methcor_df, Category == "Offenses")
meth_overdose_cor <- cor(methover$count, methoff$count)

cor_df <- overdose %>%
  spread(key = year, value = count)%>%
  gather(year, count, -Category)
cor_df$count <- ifelse(is.na(cor_df$count) == TRUE, 0, cor_df$count)

over <- filter(cor_df, Category == "Overdoses")
off <- filter(cor_df, Category == "Offenses")
heroin_overdose_cor <- cor(over$count, off$count)
```

Regardless of what *caused* the surge in heroin use, the grim reality of the increase can been seen in the plot below. This plot shows the number
of overdoses reported in the crime data. While overdoses aren't strictly a crime, they *are* reported 
in the crime statistics.  I can't be sure that the overdoses are all heroin related(for instance, meth shows a correlation of `r round(meth_overdose_cor, 2)`), but a correlation of `r round(heroin_overdose_cor, 2)` between the heroin uptick and the overdoses is *extraordinarily* coincidental
if they aren't related. 



```{r overdose_bar_plot, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE, out.width = "90%"}
overdose%>%
  filter(year >= 2007)%>%
  ggplot(aes(x = year, y = count))+
  geom_bar(stat= 'identity', aes(fill = Category), alpha = 0.5, color = 'black', position = 'dodge')+
  scale_fill_manual(values = c("#0072B2", "#D55E00"))+
  ggtitle("Heroin Offenses and Overdoses, 2007-2016")+
  labs(x = "Year", y = "Count", fill = "")+
  scale_x_discrete(limits = seq(2007, 2016, by = 1))+
  bar.theme+
  theme(
    legend.position = "bottom",
    legend.key.size = unit(4, 'mm'),
    legend.key.width = unit(4, 'mm'),
    legend.text = element_text(size = 8),
    legend.margin = unit(0, 'mm'),
    plot.title = element_text(hjust = 0.5),
    strip.background = element_rect(fill = 'white'),
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 0, vjust = .5, size = 7)
)
```

This explosion of overdoses since 2012 really drives home the true human cost of the heroin
epidemic.  I think it is crucial to remember that these are real people with real lives
who are being ravaged by addiction and, ultimately, death.


### Mapping the Epidemic

After this sobering dose of reality, I want to see exactly where this is happening within our city.
In this day and age, we hear news of heroin crimes and overdoes almost nightly on the news, but despite this, I think many people
still assume this is something that only happens in "bad" neighborhoods with bad people. This distancing thought makes it easy to put
an unpleasant piece of information out of our immediate focus.  But doing that doesn't change the reality--
heroin use is all around us today.

The first thing I wanted to do when I decided to look into heroin use was
map out the data in some way.  Luckily, R has some brilliant mapping tools that 
allow me to do just that. The finest-grained geographic area I could sub-divide the data into
turned out to be zip codes, but this works perfectly for a type of plot typically 
known as a choropleth plot.  

Using a shapefile from [this](http://data.lojic.org/datasets/jefferson-county-ky-zip-codes) website, I
grabbed the zip code boundaries for the city of Louisville and after some cleaning, aggregation, and a *ton*
of formatting(huge thanks to [Timo Grossenbacher](https://timogrossenbacher.ch/2016/12/beautiful-thematic-maps-with-ggplot2-only/)
and his brilliant post on formatting clean choropleth plots)
we can easily see the spread of heroin use through the city.

```{r heroin_choropleth_multiples, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE, out.width = "90%"}
#load zip population data for normalized plot
# editting value for 40118 manually with data from 
# https://factfinder.census.gov/faces/nav/jsf/pages/community_facts.xhtml

# load("lou_zip_pop.Rda")
# 
# extra_zip <- data.frame("40118", 9724)
# names(extra_zip) <- c("zip_code", "population")
# pop_df <- rbind(pop_df, extra_zip)
# # Putting a dummy value in for zip codes with pop == 0.
# # These are mainly PO boxes, so there is technically no population
# # But, rarely, these can have offenses tied to them. To avoid
# # Inf errors, I assign any 0 pop zip code a high value, which
# # drives down the normalized score for that zip code.
# # Not sure if this is the best way to handle it
# pop_df$population <- ifelse(pop_df$population == 0, 100000, pop_df$population)
# 
# Heroin based filtering and tidying
heroin_offenses <- crime_lou%>%
  filter(stri_detect_regex(uor_desc, "heroin"))


heroin_zip <- heroin_offenses%>%
  group_by(zip_code, year)%>%
  summarise(value = n())%>%
  rename(region = zip_code)%>%
  spread(key = year, value = value)%>%
  gather(year, value, -region)
# 
# #Dealing with NA values
# heroin_zip$value <- ifelse(is.na(heroin_zip$value) == TRUE, 0, heroin_zip$value)
# 
# 
# # Join population data to main df
# heroin_zip <- left_join(heroin_zip, pop_df, by = c('region' = 'zip_code'))
# heroin_zip$population <- ifelse(is.na(heroin_zip$population) == TRUE, 0, heroin_zip$population)
# 
# # Creating map object
# map <- readOGR("Jefferson_County_KY_ZIP_Codes.shp", verbose = FALSE)
# 
# map@data$id <- rownames(map@data)
# map_df <- tidy(map, region = 'id')
# map_df <- left_join(map_df, map@data, by = 'id')
# 
# # Joining map data to crime data
# heroin_map_df <- full_join(map_df, heroin_zip, by= c('ZIPCODE' = 'region'))
# 
# # Adding normalized(use per 10000) use data to df
# heroin_map_df <- heroin_map_df%>%
#   mutate(normalized = (value/population) * 10000)
# heroin_map_df$normalized <- ifelse(is.nan(heroin_map_df$normalized) == TRUE, 0, heroin_map_df$normalized)

heroin_map_df <- readRDS("heroin_map_df.rds.gzip")
# Base map theme
theme_map <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = "Ubuntu Regular", color = "#22211d"),
      axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      # panel.grid.minor = element_line(color = "#ebebe5", size = 0.2),
      panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
      panel.grid.minor = element_blank(),
      plot.background = element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_rect(fill = "#f5f5f2", color = NA),
      panel.border = element_blank(),
      ...
    )
}


heroin_map_0716 <- heroin_map_df%>%
  filter(year >= 2007)

no_classes <- 6
labels <- c()

quantiles <- quantile(heroin_map_0716$value, 
                      probs = seq(0, 1, length.out = no_classes + 1))
# Discrete breaks
discrete_breaks <- c(0.5,1, 4, 14, 43)
minVal <- min(heroin_map_0716$value, na.rm = T)  
maxVal <- max(heroin_map_0716$value, na.rm = T)

#form labels
labels <- c()
brks <- c(minVal, discrete_breaks, maxVal)

for(idx in 1:length(brks)){
  labels <- c(labels,round(brks[idx + 1], 2))
}

labels <- labels[1:length(labels)-1]
# new variable in dataset based on these
heroin_map_0716$brks <- cut(heroin_map_0716$value,
                               breaks = brks,
                               include.lowest = TRUE,
                               labels = labels)
brks_scale <- levels(heroin_map_0716$brks)
labels_scale <- rev(brks_scale)

extendLegendWithExtremes <- function(p, data){
  p_grob <- ggplotGrob(p)
  legend <- gtable_filter(p_grob, "guide-box")
  legend_grobs <- legend$grobs[[1]]$grobs[[1]]
  # grab the first key of legend
  legend_first_key <- gtable_filter(legend_grobs, "key-3-1-1")
  legend_first_key$widths <- unit(2, units = "cm")
  # modify its width and x properties to make it longer
  legend_first_key$grobs[[1]]$width <- unit(2, units = "cm")
  legend_first_key$grobs[[1]]$x <- unit(0.15, units = "cm")
  
  # last key of legend
  legend_last_key <- gtable_filter(legend_grobs, "key-3-6-1")
  legend_last_key$widths <- unit(2, units = "cm")
  # analogous
  legend_last_key$grobs[[1]]$width <- unit(2, units = "cm")
  legend_last_key$grobs[[1]]$x <- unit(1.02, units = "cm")
  
  # grab the last label so we can also shift its position
  legend_last_label <- gtable_filter(legend_grobs, "label-5-6")
  legend_last_label$grobs[[1]]$x <- unit(2, units = "cm")
  
  # Insert new color legend back into the combined legend
  legend_grobs$grobs[legend_grobs$layout$name == "key-3-1-1"][[1]] <- 
    legend_first_key$grobs[[1]]
  legend_grobs$grobs[legend_grobs$layout$name == "key-3-6-1"][[1]] <- 
    legend_last_key$grobs[[1]]
  legend_grobs$grobs[legend_grobs$layout$name == "label-5-6"][[1]] <- 
    legend_last_label$grobs[[1]]
  
  # finally, I need to create a new label for the minimum value 
  new_first_label <- legend_last_label$grobs[[1]]
  # have to create the variable from the input crime data, then parse it, then eval it
  new_first_label$label <- round(min(eval(parse(text = paste0(data, "$", "value"))), na.rm = T), 2)
  new_first_label$x <- unit(-0.15, units = "cm")
  new_first_label$hjust <- 1
  
  legend_grobs <- gtable_add_grob(legend_grobs, 
                                  new_first_label, 
                                  t = 6, 
                                  l = 2, 
                                  name = "label-5-0", 
                                  clip = "off")
  legend$grobs[[1]]$grobs[1][[1]] <- legend_grobs
  p_grob$grobs[p_grob$layout$name == "guide-box"][[1]] <- legend
  
  # the plot is now drawn using this grid function
  grid::grid.newpage()
  grid::grid.draw(p_grob)
}


q <- ggplot()+
  geom_polygon(data = heroin_map_0716, aes(fill = brks,
                                              x = long, 
                                              y = lat, 
                                              group = group))+
  geom_path(data = heroin_map_0716, aes(x = long,
                                           y = lat, 
                                           group = group),
            color = 'white', size = 0.1)+
  facet_wrap(~year,
             nrow = 3)+
  coord_equal()+
  theme_map()+
  theme(
    legend.background = element_rect(fill = alpha('white', 0.0)),
    legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"),
    plot.title = element_text(hjust = 0.5, color = "#4e4d47"),
    plot.subtitle = element_text(hjust = 0.5, color = "#4e4d47", 
                                 margin = margin(b = -0.1, 
                                                 t = -0.1, 
                                                 l = 2, 
                                                 unit = "cm"), 
                                 debug = F),
    legend.title = element_text(size = 8),
    plot.margin = unit(c(.5,.5,.2,.5), "cm"),
    # panel.spacing = unit(c(-.1,0.2,.2,0.2), "cm"),
    panel.border = element_blank(),
    plot.caption = element_text(size = 6,
                                hjust = 0.92,
                                margin = margin(t = 0.2,
                                                b = 0,
                                                unit = "cm"),
                                color = "#939184")
  )+
  theme(legend.position = 'bottom')+
  labs(title = "Louisville's Heroin Epidemic",
       subtitle = "Heroin Offenses by Zip Code, 2007-2016")+
  scale_fill_manual(
    # magma with 8 classes
    values = rev(magma(8, alpha = 0.7)[2:7]),
    breaks = rev(brks_scale),
    name = "Heroin Offenses",
    drop = FALSE,
    labels = labels_scale,
    guide = guide_legend(
      direction = 'horizontal',
      keyheight = unit(2, units = 'mm'),
      keywidth = unit(70/length(labels), units = 'mm'),
      title.position = 'top',
      title.hjust = 0.5,
      label.hjust = 1,
      nrow = 1,
      byrow = TRUE,
      reverse = TRUE,
      label.position = 'bottom'
    )
  )

extendLegendWithExtremes(q, "heroin_map_0716")

```

```{r greater_than, echo = FALSE, message = FALSE, cache = TRUE, warning = FALSE}
greater_than <- heroin_zip%>%
  filter(year == 2016)%>%
  mutate(greater_than = ifelse(value >=26, TRUE, FALSE))

```

The quick saturation of the city is pretty shocking. Even after seeing the sharp rise in 2012 in the plots earlier,
the near immediate darkening of the west and south end in 2012 took me by surprise. In 2016, we had 11 zip codes
that saw at least 1 heroin crime per week.  In 40203, they saw a heroin crime every `r round(365/143, 1)`
days! 

```{r biggest_change_zip_codes, echo = FALSE, message = FALSE, cache = TRUE, warning = FALSE}
changes_by_zip <- heroin_zip%>%
  # Going to add 1 to all values to work around Inf values
  mutate(value = value + 1)%>%
  filter(year == 2007|year == 2016)%>%
  spread(key = year, value = value)%>%
  mutate(per_change = `2016`/`2007` * 100)

hurstbourne <- changes_by_zip%>%
  filter(region == 40222)%>%
  dplyr::select(per_change)
```
  
But perhaps even more shocking is that even the 'better' parts of the city are being hit. The east end and regions like Prospect--
typically thought of as very affluent parts of the city-- have seen huge increases as well. We see east end regions with 
increases as high as *`r hurstbourne$per_change`%*. This is definitely not something that is only happening in one area of the city.
An incredible `r round(sum(greater_than$greater_than)/nrow(greater_than)*100, 1)`% of Louisville zip codes
are seeing at least 1 heroin crime every other week.  Keep in mind that this is heroin *crime* data--meaning that heroin
*use* is likely to be significantly higher in all regions.  I shudder to think about the true number of overdoses
or the number of families affected by the epidemic.

It should be noted that none of this really changes when you normalize for population density.  The regions with high 
counts downtown are *not* merely high because they have a higher population density. In fact, I was going to included a 
normalized map, but there was so little difference that it didn't seem to warrant the space.


```{r normalized_heroin_use, echo = FALSE, message = FALSE, cache = TRUE, warning = FALSE, eval = FALSE, out.width = "90%"}
# Not actually going to use this either. The differences over the 
# unnormalized plot don't seem to justify taking up the space
heroin_map_normalized_0716 <- heroin_map_df%>%
  filter(year >= 2007)

no_classes <- 6
labels <- c()

quantiles <- quantile(heroin_map_normalized_0716$normalized, 
                      probs = seq(0, 1, length.out = no_classes + 1))
# Discrete breaks
discrete_breaks <- c(0.5, 1, 2, 7, 14)
minVal <- min(heroin_map_normalized_0716$normalized, na.rm = T)  
maxVal <- max(heroin_map_normalized_0716$normalized, na.rm = T)

#form labels
labels <- c()
brks <- c(minVal, discrete_breaks, maxVal)

for(idx in 1:length(brks)){
  labels <- c(labels,round(brks[idx + 1], 2))
}

labels <- labels[1:length(labels)-1]
# new variable in dataset based on these
heroin_map_normalized_0716$brks <- cut(heroin_map_normalized_0716$normalized,
                               breaks = brks,
                               include.lowest = TRUE,
                               labels = labels)
brks_scale <- levels(heroin_map_normalized_0716$brks)
labels_scale <- rev(brks_scale)




q <- ggplot()+
  geom_polygon(data = heroin_map_normalized_0716, aes(fill = brks,
                                              x = long, 
                                              y = lat, 
                                              group = group))+
  geom_path(data = heroin_map_normalized_0716, aes(x = long,
                                           y = lat, 
                                           group = group),
            color = 'white', size = 0.1)+
  facet_wrap(~year,
             nrow = 3)+
  coord_equal()+
  theme_map()+
  theme(
    legend.background = element_rect(fill = alpha('white', 0.0)),
    legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"),
    plot.title = element_text(hjust = 0.5, color = "#4e4d47"),
    plot.subtitle = element_text(hjust = 0.5, color = "#4e4d47", 
                                 margin = margin(b = -0.1, 
                                                 t = -0.1, 
                                                 l = 2, 
                                                 unit = "cm"), 
                                 debug = F),
    legend.title = element_text(size = 8),
    plot.margin = unit(c(.5,.5,.2,.5), "cm"),
    # panel.spacing = unit(c(-.1,0.2,.2,0.2), "cm"),
    panel.border = element_blank(),
    plot.caption = element_text(size = 6,
                                hjust = 0.92,
                                margin = margin(t = 0.2,
                                                b = 0,
                                                unit = "cm"),
                                color = "#939184")
  )+
  theme(legend.position = 'bottom')+
  labs(title = "Louisville's Heroin Epidemic",
       subtitle = "Normalized Heroin Offenses by Zip Code, 2007-2016")+
  scale_fill_manual(
    # magma with 8 classes
    values = rev(magma(8, alpha = 0.7)[2:7]),
    breaks = rev(brks_scale),
    name = "Heroin Offenses",
    drop = FALSE,
    labels = labels_scale,
    guide = guide_legend(
      direction = 'horizontal',
      keyheight = unit(2, units = 'mm'),
      keywidth = unit(70/length(labels), units = 'mm'),
      title.position = 'top',
      title.hjust = 0.5,
      label.hjust = 1,
      nrow = 1,
      byrow = TRUE,
      reverse = TRUE,
      label.position = 'bottom'
    )
  )

extendLegendWithExtremes(q, "heroin_map_normalized_0716")


```

```{r overdose_map, echo = FALSE, message = FALSE, cache = TRUE, warning = FALSE, eval = FALSE}
# Don't think this ended up being super illustrative, so I'm not including it in the report
# Keeping here for future reference

# Overdose filtering
overdose_df <- crime_lou%>%
  filter(stri_detect_regex(uor_desc, "overdose"))


overdose_zip <- overdose_df%>%
  group_by(zip_code, year)%>%
  summarise(value = n())%>%
  rename(region = zip_code)%>%
  spread(key = year, value = value)%>%
  gather(year, value, -region)

#Dealing with NA values
overdose_zip$value <- ifelse(is.na(overdose_zip$value) == TRUE, 0, overdose_zip$value)


# Join population data to main df
overdose_zip <- left_join(overdose_zip, pop_df, by = c('region' = 'zip_code'))
overdose_zip$population <- ifelse(is.na(overdose_zip$population) == TRUE, 0, overdose_zip$population)

# Creating map object
map <- readOGR("Jefferson_County_KY_ZIP_Codes.shp", verbose = FALSE)

map@data$id <- rownames(map@data)
map_df <- tidy(map, region = 'id')
map_df <- left_join(map_df, map@data, by = 'id')

# Joining map data to crime data
overdose_map_df <- full_join(map_df, overdose_zip, by= c('ZIPCODE' = 'region'))

# Adding normalized(use per 10000) use data to df
overdose_map_df <- overdose_map_df%>%
  mutate(normalized = (value/population) * 10000)
overdose_map_df$normalized <- ifelse(is.nan(overdose_map_df$normalized) == TRUE, 0, overdose_map_df$normalized)

overdose_map_normalized_0716 <- overdose_map_df%>%
  filter(year >= 2007)

no_classes <- 6
labels <- c()

quantiles <- quantile(overdose_map_normalized_0716$normalized, 
                      probs = seq(0, 1, length.out = no_classes + 1))
# Discrete breaks
discrete_breaks <- c(0.5, 1, 2, 3, 4)
minVal <- min(overdose_map_normalized_0716$normalized, na.rm = T)  
maxVal <- max(overdose_map_normalized_0716$normalized, na.rm = T)

#form labels
labels <- c()
brks <- c(minVal, discrete_breaks, maxVal)

for(idx in 1:length(brks)){
  labels <- c(labels,round(brks[idx + 1], 2))
}

labels <- labels[1:length(labels)-1]
# new variable in dataset based on these
overdose_map_normalized_0716$brks <- cut(overdose_map_normalized_0716$normalized,
                               breaks = brks,
                               include.lowest = TRUE,
                               labels = labels)
brks_scale <- levels(overdose_map_normalized_0716$brks)
labels_scale <- rev(brks_scale)


q <- ggplot()+
  geom_polygon(data = overdose_map_normalized_0716, aes(fill = brks,
                                              x = long, 
                                              y = lat, 
                                              group = group))+
  geom_path(data = overdose_map_normalized_0716, aes(x = long,
                                           y = lat, 
                                           group = group),
            color = 'white', size = 0.1)+
  facet_wrap(~year,
             nrow = 3)+
  coord_equal()+
  theme_map()+
  theme(
    legend.background = element_rect(fill = alpha('white', 0.0)),
    legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"),
    plot.title = element_text(hjust = 0.5, color = "#4e4d47"),
    plot.subtitle = element_text(hjust = 0.5, color = "#4e4d47", 
                                 margin = margin(b = -0.1, 
                                                 t = -0.1, 
                                                 l = 2, 
                                                 unit = "cm"), 
                                 debug = F),
    legend.title = element_text(size = 8),
    plot.margin = unit(c(.5,.5,.2,.5), "cm"),
    # panel.spacing = unit(c(-.1,0.2,.2,0.2), "cm"),
    panel.border = element_blank(),
    plot.caption = element_text(size = 6,
                                hjust = 0.92,
                                margin = margin(t = 0.2,
                                                b = 0,
                                                unit = "cm"),
                                color = "#939184")
  )+
  theme(legend.position = 'bottom')+
  labs(title = "Louisville's overdose Epidemic",
       subtitle = "Normalized overdose Offenses by Zip Code, 2007-2016")+
  scale_fill_manual(
    # magma with 8 classes
    values = rev(magma(8, alpha = 0.7)[2:7]),
    breaks = rev(brks_scale),
    name = "overdose Offenses",
    drop = FALSE,
    labels = labels_scale,
    guide = guide_legend(
      direction = 'horizontal',
      keyheight = unit(2, units = 'mm'),
      keywidth = unit(70/length(labels), units = 'mm'),
      title.position = 'top',
      title.hjust = 0.5,
      label.hjust = 1,
      nrow = 1,
      byrow = TRUE,
      reverse = TRUE,
      label.position = 'bottom'
    )
  )

extendLegendWithExtremes(q, "overdose_map_normalized_0716")


```

### Conclusion
So, sadly, it seems since around 2012 Louisville has been in the midst of a heroin epidemic. If you don't think it's happening near you,
you are being blinded by wishful thinking.  Heroin crime--and one can then surmise heroin use-- is happening all around us.  While I don't
proclaim to know the solution to such a complex, multifaceted problem, I *do* think the first step is to acknowledgement that
this is an issue for everyone in our community.  The consequences of this epidemic--be it in lives taken, or families shattered,
or property destroyed--are affecting people of all ages, races, genders and incomes.

As usual, all code used in this report can be found [here](https://github.com/Grollus/DataBlog/tree/master/content/post).